---
title: "Text Analysis"
site: distill::distill_website
output: distill::distill_article
---
```{r include=FALSE}
load("run_unbanked.RData")
library(ggplot2)
library(dplyr)
library(tidytext)
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

## Data Preparation
```{r}

text_data <- run_unbanked %>%
  select(Q40_9_TEXT) %>%
  na.omit()

words <- text_data %>%
  unnest_tokens(word, Q40_9_TEXT) %>%
  count(word, sort = TRUE)

head(words, 10)
```

## Top Word Frequencies
```{r}
words %>%
  top_n(10, n) %>%
  ggplot(aes(x = reorder(word, n), y = n)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Most common words", x = "Word", y = "Count") +
  theme_minimal()
```

## Text Analysis Example

I conducted a simple word frequency analysis on open-ended survey responses.

The most common words were "time", "full", and "work", suggesting that respondents frequently mention time constraints and employment-related issues.

This indicates that themes such as work, education, and caregiving responsibilities are central in the data.
